---
title: "Figures for 3A"
author: "Frederick Boehm"
date: "11/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

First, we load the Keller 2018 data.

```{r}
load("../data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
readRDS("../data/hotspot_expr_tib2_keller_chr2.rds") -> hotspot_expr
readRDS("../data/keller2018-chr2-local-expr.rds") -> local_expr
```

```{r}
names_21 <- colnames(local_expr)[-22]
```

We isolate the names of the 21 local genes that we've been studying in 3A. 

Now, let's recreate the `local_expr` object from the Attie file's data.

```{r}
d1 <- dataset.islet.rnaseq$lod.peaks %>%
  as_tibble() %>%
  filter(chrom == 2) %>%
  left_join(dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(lod > 40) %>%
  filter(middle > 165.5 - 2, middle < 165.5 + 2, pos > 165.5 - 2, pos < 165.5 + 2) 
```

In examining the Rmd file "2018-09-10_condor-setup.Rmd" we see the object `trait_names` is defined and it's the genes in `trait_names` that are used in our previous analyses.

I wish to amend this list of genes. First, there are three of 21 genes that are way too far from the hotspot to be useful. Second, I want criteria that are easier to explain. 

We'll use the criteria that we wrote above to define d1. Note that it has 19 genes. 

So, 

1. LOD > 40
2. middle within 5 Mb of hotspot center (165.5)
3. QTL peak within 5 Mb of hotspot center (165.5)

Now, let's isolate the genes that are in d1 but not in names_21.

```{r}
d1$annot.id
d1$annot.id[!(d1$annot.id %in% names_21)]
```

We see that two traits are not in `names_21`. 

Let's get the expression data for these two and save it as a RDS file.

```{r}
dataset.islet.rnaseq$expr[ , colnames(dataset.islet.rnaseq$expr) %in% d1$annot.id[!(d1$annot.id %in% names_21)]] %>%
  saveRDS(file = "../data/keller2018-chr2-local-expr-2-2018-11-29.rds")
```

## Update on Dec 9, 2018

We need to analyze an extra 8 trans genes. To facilitate this, I'll create a "local" genes expression file for use with Condor.

```{r}
local_expr13 <- dataset.islet.rnaseq$expr[ , colnames(dataset.islet.rnaseq$expr) %in% d1$annot.id]
saveRDS(local_expr13, file = "../data/local_expr13.rds")
```



## Session info

```{r}
devtools::session_info()
```


