---
title: "Identifying Chr 2 hotspot genes without 2d scans"
author: "Frederick Boehm"
date: "12/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Our goal is to identify genes that 1) map to Chr 2 hotspot in the Keller 2018 data and 2) weren't included in my earlier 2-dimensional scans.

First, we load the csv file (which I created with code in the file "2018-12-05_FileS2Attie_eQTL_paper_eQTL_hotspots.Rmd") that contains the names of the 147 genes that meet the criteria for mapping to Chr 2 hotspot.

```{r}
library(tidyverse)
```

```{r}
read_csv("chr2_hotspot_genes.csv") -> gatti_chr2
```

We see that the genes are identified by gene symbol in the object `gatti_chr2`.

Now, read the rds file that contains the genes that I've analyzed previously.

```{r}
hot_expr <- readRDS("../data/hotspot_expr_tib2_keller_chr2.rds")
colnames(hot_expr)[-140] -> genes_139
```

One question I've had is: Why did I fail to analyze some Chr2 hotspot genes in my earlier analysis? 

The answer: because some of them are not annotated as chr2 hotspot genes.

We see this by tabulating the hotspot variable in `gatti_chr2`:

```{r}
table(gatti_chr2$hotspot)
```

Let's subset the dataframe to only those genes that have hotspot == chr5 or chr7.

```{r}
gatti_chr2 %>%
  filter(hotspot == "chr5" | hotspot == "chr7")
gatti_chr2a <- dataset.islet.rnaseq$annots %>%
  select(gene_id, symbol) %>%
  right_join(gatti_chr2) %>%
  filter(hotspot == "chr5" | hotspot == "chr7")
```


## Make the expression data frame for the 'missing' genes

We now need to prepare an expression dataframe for the 8 "missing" genes.

```{r}
expr_8 <- dataset.islet.rnaseq$expr[ , colnames(dataset.islet.rnaseq$expr) %in% gatti_chr2a$gene_id]
```

```{r}
saveRDS(expr_8, file = "../data/hotspot-chr2-8missing-genes.rds")
```

