---
title: "Comparing LRT for pleiotropy v separate QTL against LOD diff in mediation analyses"
author: "Frederick Boehm"
date: "`r lubridate::now()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

Our goal here is to examine the relationships between pleiotropy v separate QTL analyses and mediation analyses. We study with these methods the chromosome 2 hotspot in the Keller et al 2018 pancreatic islet expression data.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

First, we load the Keller 2018 data.

```{r}
load("../data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
readRDS("../data/hotspot_expr_tib2_keller_chr2.rds") -> hot139
readRDS("../data/hotspot-chr2-8missing-genes.rds") -> hot8
readRDS("../data/local_expr13.rds") -> local13
cbind(hot8, as.matrix(hot139[, -140])) -> hot
```

We now have a 378 by 13 matrix of local expression values, ie, with the updated collection of local genes.

We next need the scan1 outputs for these traits.

```{r}
library(qtl2)
```



```{r}
# define addcovar 
addcovar <- dataset.islet.rnaseq$covar
# scan1 calls
if(!file.exists("../data/s1_local.rds")){
  scan1(genoprobs = genoprobs, pheno = local13, kinship = K, addcovar = addcovar, reml = TRUE, cores = 6
        ) -> s1_local
  saveRDS(s1_local, file = "../data/s1_local.rds")
}

if(!file.exists("../data/s1_hotspot.rds")){
  scan1(genoprobs = genoprobs, pheno = hot, kinship = K, addcovar = addcovar, reml = TRUE, cores = 6
        ) -> s1_hotspot
  saveRDS(s1_hotspot, "../data/s1_hotspot.rds")
}
```





```{r}
# what is the ENSEMBL id for Hnf4a?
hnf4a_id <- "ENSMUSG00000017950"
```

We still need to define the driver for input to `intermediate::mediation_scan`.


## Load scan1() outputs for hotspot (ie, trans) traits

We saved the scan1 outputs, because they take some time to re-run.

```{r}
s1_local <- readRDS("../data/s1_local.rds")
s1_hotspot <- readRDS("../data/s1_hotspot.rds")
```

## Find QTL peaks

We now search the outputs of `scan1` to identify peaks:

```{r}
# for local genes, we above defined those of interest to have 
# peak > 40, so we use 40 here as threshold
find_peaks(s1_local, map = map, threshold = 40) -> local_peaks
find_peaks(s1_hotspot, map = map) -> hotspot_peaks
```

```{r}
map_chr2_tib <- tibble(pos = map$`2`, index = 1:length(map$`2`))
hnf4a_peak_position <- local_peaks %>%
  filter(lodcolumn == hnf4a_id) %>% 
  left_join(map_chr2_tib, by = "pos") %>%
  select(index) %>% 
  unlist() %>%
  as.numeric()
```



Let's subset the driver and the covariates:

```{r}
addcovar2 <- addcovar[rownames(addcovar) %in% rownames(hot), ]
```




```{r}
rna_annot <- dataset.islet.rnaseq$annots %>%
  as_tibble() %>%
  select(gene_id, symbol)
```


## Set up mediation analyses for all 13 local genes

First, we want to set up the `dri21` driver object. To do that, we need to know the peak positions for each local trait.

```{r}
library(intermediate) # from github, fboehm/intermediate
```



```{r}
indices <- local_peaks %>%
  left_join(map_chr2_tib) %>%
  select(index) %>%
  unlist()
```

```{r}
dri21 <- genoprobs$`2`[rownames(genoprobs$`2`) %in% rownames(hot), , indices]
cbind(dimnames(dri21)[[3]], 
indices)
```
```{r}
med21 <- local13[ , colnames(local13) %in% local_peaks$lodcolumn]
med21 %>% colnames()
```





```{r}
out21 <- list()
foo <- list()
for (j in 1:13){
  for (i in 1:147){
   foo[[i]]<- fit1_med(driver = dri21[ , , j],
         target = hot[ , i], 
         mediator = med21[ , j], 
         addcovar = addcovar2)
  }
  out21[[j]] <- foo
}
```

```{r}
med_out <- out21 %>%
  purrr::map(bind_rows)
m2_out <- med_out %>%
  bind_rows() %>%
  as_tibble() %>%
  mutate(local_gene_id = rep(colnames(med21), each = 147), nonlocal_gene_id = rep(colnames(hot), times = 13)) %>%
  mutate(lod_without_med = (ll_d - ll_1) / log(10), lod_with_med = (ll_dm - ll_1m) / log(10), lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med)

```





## Tabulating the pleiotropy v separate QTL LRT statistics

Our goal is to analyze the results of pvl scan results files from CHTC.

```{r}
library(qtl2chtc)
```

```{r}
dir(path = "../results/pvl-run2001/", pattern = "*.txt") -> fns
  lrt <- numeric()
  i <- 1
  for (fn in fns){
    read.table(file.path("../results/pvl-run2001/", fn)) -> scan_out
    qtl2pleio::calc_lrt_tib(scan_out) -> lrt[i]
    i <- i + 1
  }
tibble::tibble(fns, lrt) -> r2001
```


Let's parse the file names in `r2001` into meaningful fields, like gene names.

```{r}
library(stringr)
```

```{r}
splitted <- r2001$fns %>% 
  str_split(pattern = "_") 

my_split <- function(string_vec, pattern = ".txt"){
  foo <- str_split(string_vec, pattern)
  sapply(FUN = function(x)x[1], X = foo)
}
res2001 <- sapply(FUN = function(x)x[1:4], X = splitted) %>% 
  t() %>%
  as_tibble() %>%
  rename(run_num = V1, job_num = V2, local_gene_id = V3) %>%
  mutate(nonlocal_gene_id = my_split(V4)) %>%
  select(- V4) %>%
  mutate(lrt = r2001$lrt) %>%
  arrange(desc(lrt)) %>%
  left_join(rna_annot, by = c("nonlocal_gene_id" = "gene_id")) %>%
  rename(nonlocal_gene_symbol = symbol) %>%
  left_join(rna_annot, by = c("local_gene_id" = "gene_id")) %>%
  rename(local_gene_symbol = symbol)
```


```{r}
hist(res2001$lrt)
```

```{r}
sum(res2001$lrt > 4)
```



## Join pleiotropy v separate QTL LRT tibble with LOD drops tibble

```{r}
library(plotly)
```
    
```{r}
myt2 <- m2_out %>% 
  left_join(res2001) 
pp <- myt2 %>%
  ggplot() + aes(x = lrt, y = lod_diff, colour = nonlocal_gene_symbol) + geom_point() + theme(legend.position="none")
ggplotly(pp, tooltip = c("colour", "x", "y")) 
```


```{r}
pp <- myt2 %>%
  ggplot() + aes(x = lrt, y = lod_diff_proportion, colour = nonlocal_gene_symbol) + geom_point() + theme(legend.position="none")
ggplotly(pp, tooltip = c("colour", "x", "y")) 

```

## Get LRTs for pleiotropy v separate QTL for all 13 local genes

### run 2002: tabulating lrt statistics

```{r}
dir(path = "../results/pvl-run2002/", pattern = "*.txt") -> fns
lrt <- numeric()
i <- 1
for (fn in fns){
    read.table(file.path("../results/pvl-run2002/", fn)) -> scan_out
    qtl2pleio::calc_lrt_tib(scan_out) -> lrt[i]
    i <- i + 1
}
tibble::tibble(fns, lrt) -> r2002
```

```{r}
res2002 <- r2002 %>%
  select(fns) %>%
  unlist() %>%
  str_split_fixed(pattern = "_", n = 4) %>%
  as_tibble() %>%
  rename(run = V1, job = V2, local_gene_id = V3, nonlocal_gene_id = V4) %>%
  mutate(job = as.numeric(job)) %>%
  mutate(nonlocal_gene_id = str_remove(nonlocal_gene_id, pattern = ".txt")) %>%
  mutate(lrt = r2002$lrt) %>%
  arrange(desc(lrt)) %>%
  left_join(rna_annot, by = c("nonlocal_gene_id" = "gene_id")) %>%
  rename(nonlocal_gene_symbol = symbol) %>%
  left_join(rna_annot, by = c("local_gene_id" = "gene_id")) %>%
  rename(local_gene_symbol = symbol)
```

### Get test statistics for run 2003

```{r}
dir(path = "../results/pvl-run2003/", pattern = "*.txt") -> fns
lrt <- numeric()
i <- 1
for (fn in fns){
    read.table(file.path("../results/pvl-run2003/", fn)) -> scan_out
    qtl2pleio::calc_lrt_tib(scan_out) -> lrt[i]
    i <- i + 1
}
tibble::tibble(fns, lrt) -> r2003
```

```{r}
res2003 <- r2003 %>%
  select(fns) %>%
  unlist() %>%
  str_split_fixed(pattern = "_", n = 4) %>%
  as_tibble() %>%
  rename(run = V1, job = V2, local_gene_id = V3, nonlocal_gene_id = V4) %>%
  mutate(job = as.numeric(job)) %>%
  mutate(nonlocal_gene_id = str_remove(nonlocal_gene_id, pattern = ".txt")) %>%
  mutate(lrt = r2003$lrt) %>%
  arrange(desc(lrt)) %>%
  left_join(rna_annot, by = c("nonlocal_gene_id" = "gene_id")) %>%
  rename(nonlocal_gene_symbol = symbol) %>%
  left_join(rna_annot, by = c("local_gene_id" = "gene_id")) %>%
  rename(local_gene_symbol = symbol)
```



### Get test statistics for run 2004

```{r}
dir(path = "../results/pvl-run2004/", pattern = "*.txt") -> fns
lrt <- numeric()
i <- 1
for (fn in fns){
    read.table(file.path("../results/pvl-run2004/", fn)) -> scan_out
    qtl2pleio::calc_lrt_tib(scan_out) -> lrt[i]
    i <- i + 1
}
tibble::tibble(fns, lrt) -> r2004
```

```{r}
res2004 <- r2004 %>%
  select(fns) %>%
  unlist() %>%
  str_split_fixed(pattern = "_", n = 4) %>%
  as_tibble() %>%
  rename(run = V1, job = V2, local_gene_id = V3, nonlocal_gene_id = V4) %>%
  mutate(job = as.numeric(job)) %>%
  mutate(nonlocal_gene_id = str_remove(nonlocal_gene_id, pattern = ".txt")) %>%
  mutate(lrt = r2004$lrt) %>%
  arrange(desc(lrt)) %>%
  left_join(rna_annot, by = c("nonlocal_gene_id" = "gene_id")) %>%
  rename(nonlocal_gene_symbol = symbol) %>%
  left_join(rna_annot, by = c("local_gene_id" = "gene_id")) %>%
  rename(local_gene_symbol = symbol)
```





We then combine `res2001`, `res2002`, `res2003`, and `res2004`.

```{r}
res <- bind_rows(res2001, res2002, res2003, res2004)
```

```{r}
fn <- "../data/res-tibble-lrt-med.rds"
if (!(file.exists(fn))){
  saveRDS(res, fn)
}
readRDS(fn) -> res
```


## Work with all 13 * 147 = 1911 test statistics & mediation analyses




```{r}
myt3 <- m2_out %>% 
  left_join(res) 

pp <- myt3 %>%
  ggplot() + aes(x = lrt, y = lod_diff_proportion * (lod_diff_proportion > 0), colour = local_gene_symbol == "Hnf4a") + geom_point(size = 1 / 10) + broman::karl_theme() +  theme(legend.position="none") + ylab("LOD difference proportion") + xlab("Pleiotropy vs. separate QTL test statistic") + ylim(0, 1)
ggplotly(pp, tooltip = c("colour", "x", "y")) 
ggsave("lod-diff-prop-v-lrt.jpg", height = 7, width = 7, units = "in")
ggsave("lod-diff-prop-v-lrt.eps", height = 7, width = 7, units = "in")
ggsave("lod-diff-prop-v-lrt.svg", height = 7, width = 7, units = "in")
```

## Repeat above plot, but with only 147 pairs involving *Hnf4a*

```{r}
pp <- myt3 %>%
  filter(local_gene_symbol == "Hnf4a") %>%
  ggplot() + aes(x = lrt, y = lod_diff_proportion * (lod_diff_proportion > 0), colour = local_gene_symbol) + geom_point(size = 1 / 4) + broman::karl_theme() + theme(legend.position="none") + ylab("LOD difference proportion") + xlab("Pleiotropy vs. separate QTL test statistic") + ylim(0, 1)
ggplotly(pp, tooltip = c("colour", "x", "y")) 
ggsave("Hnf4a-lod-diff-prop-v-lrt.jpg", height = 7, width = 7, units = "in")
ggsave("Hnf4a-lod-diff-prop-v-lrt.eps", height = 7, width = 7, units = "in")
ggsave("Hnf4a-lod-diff-prop-v-lrt.svg", height = 7, width = 7, units = "in")
```




```{r}
foo <- m2_out %>% 
  left_join(res) %>%
  filter(!(local_gene_symbol == "Hnf4a")) 
unique(foo$local_gene_symbol) -> local_unique
# find max value of lrt in the full df
max(foo$lrt)+ 1 -> xmax
# separate plots for each local gene
for (i in 1:12){
  foo %>%
    filter(local_gene_symbol == local_unique[i]) %>%
    ggplot() + geom_point(aes(x = lrt, y = lod_diff_proportion * (lod_diff_proportion > 0)), size = 1/10) + broman::karl_theme(strip.placement = "inside") + theme(legend.position="none") + ylab("LOD difference proportion") + xlab("Pleiotropy vs. separate QTL test statistic") + ylim(0, 1) + xlim(0, xmax)
  ggsave(paste0("bar_", i, ".jpg"), height = 3, width = 6, units = "in")
}
```





```{r}
myt4 <- m2_out %>% 
  left_join(res) 
```

Now, a figure with truncated lod diff proportion values.

```{r}
unique(myt4$local_gene_symbol) -> local_symbols
local_symbols
local_symbols12 <- local_symbols[-4] # remove Hnf4a
jpeg("baseR-12.jpg", height = 720)
par(mfrow = c(4, 3), mar = c(4, 4, 1, 1))
for (i in 1:12){
  myt4_sub <- myt4 %>% 
    filter(local_gene_symbol == local_symbols12[i]) 
  plot(y = myt4_sub$lod_diff_proportion * (myt4_sub$lod_diff_proportion > 0), 
       x = myt4_sub$lrt, 
       ylab = "LOD difference proportion", 
       xlab = "Pleiotropy vs. separate QTL test statistic",
       ylim = c(0, 1),
       xlim = c(0, 73),
       cex.lab = 0.7, # reduce size of labels
       cex.axis = 0.7, # 
       mgp = c(3, 2, 1) # default is (3, 1, 0). 
       )
  title(myt4_sub$local_gene_symbol[1], adj = 0.9, line = -2)
}
dev.off()
```

Now, the analogous plot for Hnf4a. 

```{r}
jpeg("baseR-Hnf4a.jpg")
myt4_sub <- myt4 %>% 
  filter(local_gene_symbol == "Hnf4a") 
plot(y = myt4_sub$lod_diff_proportion * (myt4_sub$lod_diff_proportion > 0), 
      x = myt4_sub$lrt, 
      ylab = "LOD difference proportion", 
      xlab = "Pleiotropy vs. separate QTL test statistic",
      ylim = c(0, 1),
      xlim = c(0, 73),
)
title("Hnf4a", adj = 0.9, line = -2)
dev.off()

```





Note that Gpr20 has a high LRT for pleiotropy v separate QTL with Hnf4a, yet it has a small LRT when paired with Ptgis. In both cases, the proportion diff in LOD is very small. 

Let's look at the table of results that involve Gpr20

```{r}
m2_out %>% 
  left_join(res) %>%
  filter(nonlocal_gene_symbol == "Gpr20") %>%
  arrange(desc(lod_diff_proportion))
```
It looks like Gpr20 has small LOD scores for most loci examined here.

```{r}
gpr20_out <- m2_out %>% 
  left_join(res) %>%
  filter(nonlocal_gene_symbol == "Gpr20") %>%
  arrange(desc(lod_without_med))
```

```{r}
symbols <- dataset.islet.rnaseq$annots %>%
  as_tibble() %>%
  select(gene_id, symbol)
local2 <- local_peaks %>%
  as_tibble() %>%
  left_join(symbols, by = c("lodcolumn" = "gene_id")) %>%
  rename(local_gene_lod = lod, local_gene_id = lodcolumn)
# Get position of Gpr20's peak on chr2
Gpr20_pos <- dataset.islet.rnaseq$lod.peaks %>%
  as_tibble() %>%
  filter(chrom == 2) %>%
  filter(annot.id == symbols$gene_id[which(symbols$symbol == "Gpr20")]) %>%
  select(pos) %>% unlist()


pp <- local2 %>%
  left_join(gpr20_out) %>%
  tidyr::gather(lod_diff_proportion, lrt, key = "method", value = "test_stat") %>%
  ggplot() + geom_point(aes(x = pos, y = test_stat, colour = local_gene_symbol)) + facet_wrap(~method, ncol = 1, scales="free") + geom_vline(xintercept = Gpr20_pos)
# add vertical line for nonlocal gene's univariate QTL peak?   
ggplotly(pp, tooltip = c("colour", "x", "y"))  
```



## How many nonlocal genes have Hnf4a as the strongest mediator?

```{r}
library(xtable)
```




```{r}
m2_out %>% 
  left_join(res) %>%
  left_join(local2) %>%
  group_by(nonlocal_gene_symbol) %>%
  filter(lod_diff == max(lod_diff)) %>%
  ungroup() %>%
  select(local_gene_symbol) %>%
  table() %>% xtable() %>% print.xtable()
```

## Per-nonlocal gene analyses

Here, our goal is to create 147 individual plots. Each plot will have two panes, stacked vertically. The top pane is for chromosome position (of local gene's middle, x axis) and y axis is pleiotropy v separate QTL test statistic. The bottom pane is for chromosome position (of local gene's middle) and, on the y axis, LOD difference proportion.

```{r}
# first, join in the peak positions from hotspot_peaks
hot_peaks2 <- hotspot_peaks %>%
  filter(chr == 2)
## for use with geom_blank()
blank_data <- tibble(group = c("lrt", "lrt", "lod_diff_proportion", "lod_diff_proportion"), x = c(163, 168, 163, 168), y = c(0, 
    NA, 0, 1))
# i found this blog post for using geom_blank: https://chrischizinski.github.io/rstats/using_geom_blank/


unique(myt4$nonlocal_gene_symbol) -> nonlocal_symbols
## plots
for (i in (1:147)){
  myt4 %>%
    filter(nonlocal_gene_symbol == nonlocal_symbols[i]) %>%
    left_join(dataset.islet.rnaseq$annots, by = c("local_gene_id" = "gene_id")) %>%
    left_join(hot_peaks2, by = c("nonlocal_gene_id" = "lodcolumn")) %>%
    gather(stat, value, lrt, lod_diff_proportion) %>%
    mutate(val2 = value * (value > 0)) %>%
    ggplot() + geom_point(aes(x = middle, y = val2)) + geom_blank(data = blank_data, aes(x = x, y = y)) + geom_vline(aes(xintercept = pos)) + broman::karl_theme() + theme(legend.position="none") + ylab("Statistic") + xlab("Chromosome 2 position") + facet_wrap(nrow = 2, facets = vars(stat), shrink = FALSE, scales = "free")
  fn <- paste0("nonlocal_", i, "_chr2position")
  ggsave(paste0(fn, ".jpg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".svg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".eps"), height = 7, width = 7, units = "in")
}
```

## Separate plots instead of 2-panel plots

```{r}
for (i in (1:147)){
  foo <- myt4 %>%
    filter(nonlocal_gene_symbol == nonlocal_symbols[i]) %>%
    left_join(dataset.islet.rnaseq$annots, by = c("local_gene_id" = "gene_id")) %>%
    left_join(hot_peaks2, by = c("nonlocal_gene_id" = "lodcolumn")) %>%     
    mutate(hnf4a_indic = (local_gene_symbol == "Hnf4a"))

  # LOD diff proportion plots
  foo %>% 
    ggplot() + geom_point(aes(x = middle, y = (lod_diff_proportion > 0) * lod_diff_proportion, color = hnf4a_indic)) + geom_vline(aes(xintercept = pos)) + broman::karl_theme() + theme(legend.position="none") + ylab("LOD difference proportion") + xlab("Chromosome 2 position") + ylim(c(0, 1))
  fn <- paste0("1-panel-LOD-diff-nonlocal_", i, "_chr2position")
  ggsave(paste0(fn, ".jpg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".svg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".eps"), height = 7, width = 7, units = "in")
  # pleio LRT plots
  foo %>% 
    ggplot() + geom_point(aes(x = middle, y = lrt, color = hnf4a_indic)) + geom_vline(aes(xintercept = pos)) + broman::karl_theme() + theme(legend.position="none") + ylab("Pleiotropy v separate QTL test statistic") + xlab("Chromosome 2 position") + ylim(c(0, NA))
  fn <- paste0("1-panel-LRT-nonlocal_", i, "_chr2position")
  ggsave(paste0(fn, ".jpg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".svg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".eps"), height = 7, width = 7, units = "in")
}
```

## Scatter plots, without chromosome position data

Here, we'll create per-nonlocal gene scatter plots, but without consideration of the chromosome position. That is, it's just a scatter plot of LOD difference proportion against pleiotropy vs. separate QTL test statistic.


```{r}
for (i in (1:147)){
  foo <- myt4 %>%
    filter(nonlocal_gene_symbol == nonlocal_symbols[i]) %>%
    left_join(dataset.islet.rnaseq$annots, by = c("local_gene_id" = "gene_id")) %>%
    left_join(hot_peaks2, by = c("nonlocal_gene_id" = "lodcolumn")) %>%     
    mutate(hnf4a_indic = (local_gene_symbol == "Hnf4a"))

  # scatter plots
  foo %>% 
    ggplot() + geom_point(aes(x = lrt, y = (lod_diff_proportion > 0) * lod_diff_proportion, color = hnf4a_indic)) + broman::karl_theme() + theme(legend.position="none") + ylab("LOD difference proportion") + xlab("Pleiotropy vs. separate QTL test statistic") + ylim(c(0, 1)) + xlim(0, 73) + ggtitle(nonlocal_symbols[i]) + theme(plot.title = element_text(hjust = 0.5, vjust = - 10))
  fn <- paste0("nonlocal-scatter_", i)
  ggsave(paste0(fn, ".jpg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".svg"), height = 7, width = 7, units = "in")
  ggsave(paste0(fn, ".eps"), height = 7, width = 7, units = "in")
}
```

```{r}
# find examples for thesis images
# first, get the symbols for those with large Hnf4a test statistics
myt4 %>%
  filter(local_gene_symbol == "Hnf4a") %>%
  select(lrt) %>% 
  mutate(lrt_big_indic = lrt > 5) %>% 
  select(lrt_big_indic) %>%
  unlist() %>%
  sum()
```
We see that 14 nonlocal genes, when paired with Hnf4a, have a pleiotropy v separate QTL test statistic above 5. 

Now, let's get the 14 nonlocal gene symbols.


```{r}
nonlocal_big_lrt <- myt4 %>%
  filter(local_gene_symbol == "Hnf4a") %>%
  filter(lrt > 5) %>%
  select(nonlocal_gene_symbol) %>%
  unlist() 
```






## Counts per local gene of strongest mediator


```{r}
myt4 %>%
  filter(local_gene_symbol == "Hnf4a") %>%
  mutate(lod_med_indic = lod_diff > 1.5) %>%
  select(lod_med_indic) %>% unlist() %>% table()
```

We see that 88 of 147 genes meet the criteria for mediation by Hnf4a. This is the same count that Keller et al report in their GENETICS 2018 article. I haven't yet established whether the 88 gene ids are the same as those that Keller identified.



## xtables for 3A

We want to make two latex tables for use in the 3A manuscript. One for annotating the local genes, one for the nonlocal genes.

```{r}
library(xtable)
```

```{r}
local_annot <- local_peaks %>%
  left_join(dataset.islet.rnaseq$annots, by = c("lodcolumn" = "gene_id")) %>%
  select(lodcolumn, chr.y, pos, symbol, start, end) %>%
  rename(ensembl_id = lodcolumn, chromosome = chr.y, LOD_peak_position = pos) %>%
  select(symbol, ensembl_id, chromosome, start, end, LOD_peak_position) %>%
  arrange(LOD_peak_position) %>%
  left_join(dataset.islet.rnaseq$lod.peaks, by = c("ensembl_id"= "annot.id")) %>%
  filter(chrom == 2) %>%
  select(- pos, - marker.id, - chrom) %>%
  rename(LOD_peak_height = lod)
```

```{r}
print.xtable(xtable(local_annot), include.rownames = FALSE)
```

```{r}
local_annot2 <- local_annot %>%
  select(- ensembl_id, - chromosome)
print.xtable(xtable(local_annot2), include.rownames = FALSE)
```

## Table for nonlocal gene annotations

```{r}
hot_peaks2 %>%
  select(- lodindex, - chr) %>%
  rename(gene_id = lodcolumn, position = pos, LOD = lod) %>%
  left_join(dataset.islet.rnaseq$annots, by = "gene_id") %>%
  select(symbol, position, LOD) %>%
  arrange(position) %>%
  xtable() %>%
  print.xtable(include.rownames = FALSE, tabular.environment = "longtable")
```

## Fixing 'strip' for local & nonlocal gene plots


```{r}

foo <- m2_out %>% 
  left_join(res) %>%
  filter(!(local_gene_symbol == "Hnf4a")) %>%
  mutate(x = 60, y = 0.8)
  
unique(foo$local_gene_symbol) -> local_unique
# find max value of lrt in the full df
# separate plots for each local gene
foo %>%
  ggplot() + geom_point(aes(x = lrt, y = lod_diff_proportion * (lod_diff_proportion > 0)), size = 1/10) + broman::karl_theme() + facet_wrap(~ local_gene_symbol, nrow = 6, ncol = 2) + theme(legend.position="none") + #theme(
#  strip.background = element_blank(),
#  strip.text.x = element_blank()
#) + 
  ylab("LOD difference proportion") +
  xlab("Pleiotropy vs. separate QTL test statistic") + 
  ylim(0, 1) + xlim(0, NA) +
  geom_text(data = foo, aes(x, y, label = local_gene_symbol), inherit.aes=FALSE)

ggsave("12local-facet_grid.eps", width = 6.5, height = 9, units = "in")
```




```{r}


# separate plots for each local gene
foo %>%
  ggplot() + geom_point(aes(x = lrt, y = lod_diff_proportion * (lod_diff_proportion > 0)), size = 1/10) + broman::karl_theme() + facet_wrap(~ local_gene_symbol, nrow = 6, ncol = 2) + theme(legend.position="none") + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) + 
  ylab("LOD difference proportion") +
  xlab("Pleiotropy vs. separate QTL test statistic") + 
  ylim(0, 1) + xlim(0, NA) +
  geom_text(data = foo, aes(x, y, label = local_gene_symbol), inherit.aes=FALSE)

ggsave("12local-facet_grid-no-strip.eps", width = 6.5, height = 9, units = "in")
# https://stackoverflow.com/questions/10547487/remove-facet-wrap-labels-completely
```

Now, the nonlocal gene plots.


```{r}
indices <- c(97, 105, 64, 141)
nonlocal_symbols[indices]
levels(as.factor(nonlocal_symbols[indices]))
levs <- levels(as.factor(nonlocal_symbols[indices]))[c(1, 3, 4, 2)]
levs_logical <- c(TRUE, FALSE)
foo <- myt4 %>%
    filter(nonlocal_gene_symbol %in% nonlocal_symbols[indices]) %>%
    left_join(dataset.islet.rnaseq$annots, by = c("local_gene_id" = "gene_id")) %>%
    left_join(hot_peaks2, by = c("nonlocal_gene_id" = "lodcolumn")) %>%     
    mutate(hnf4a_indic = (local_gene_symbol == "Hnf4a"), 
           ordering_factor = factor(nonlocal_gene_symbol, levels = levs),
           hnf4a_indic_factor = factor(hnf4a_indic, levels = levs_logical), 
           x = 60, y = 0.8 # for adding geom_text
             )

  # scatter plots 

foo %>% 
  ggplot() + 
  geom_point(size = 1 + foo$hnf4a_indic, aes( 
                           x = lrt, 
                           y = (lod_diff_proportion > 0) * lod_diff_proportion, 
                           color = hnf4a_indic_factor)) + 
  #geom_jitter() + 
  broman::karl_theme() + 
  theme(legend.position="none") + 
  ylab("LOD difference proportion") + 
  xlab("Pleiotropy vs. separate QTL test statistic") + 
  ylim(c(0, 1)) + 
  xlim(0, 73) + 
  facet_wrap(~ ordering_factor, nrow = 2, ncol = 2) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  geom_text(data = foo, aes(x, y, label = nonlocal_gene_symbol), inherit.aes=FALSE)
# 
fn <- "nonlocal-4-panels-no-strip"
ggsave(paste0(fn, ".svg"), height = 9, width = 6.5, units = "in")
ggsave(paste0(fn, ".eps"), height = 9, width = 6.5, units = "in")
foo %>% 
  ggplot() + 
  geom_point(size = 1 + foo$hnf4a_indic, aes( 
                           x = lrt, 
                           y = (lod_diff_proportion > 0) * lod_diff_proportion, 
                           color = hnf4a_indic_factor)) + 
  #geom_jitter() + 
  broman::karl_theme() + 
  theme(legend.position="none") + 
  ylab("LOD difference proportion") + 
  xlab("Pleiotropy vs. separate QTL test statistic") + 
  ylim(c(0, 1)) + 
  xlim(0, 73) + 
  facet_wrap(~ ordering_factor, nrow = 2, ncol = 2) + 
#  theme(
#    strip.background = element_blank(),
#    strip.text.x = element_blank()
#) +
  geom_text(data = foo, aes(x, y, label = nonlocal_gene_symbol), inherit.aes=FALSE)
# 
fn <- "nonlocal-4-panels"
ggsave(paste0(fn, ".svg"), height = 9, width = 6.5, units = "in")
ggsave(paste0(fn, ".eps"), height = 9, width = 6.5, units = "in")

```





## Session info

```{r}
devtools::session_info()
```
